Predicting the effect of candidate interventions for mitigating disease spread in asymmetric-mixing metapopulations

### 2024-07-30 ms.tex
started analysis sketch
updated some figs

To quantitatively compare hypothetical intervention strategies we analyze reduction
in total Once implemented,

@@

@@

\begin{itemize}
  \item 
    Application to other problematic diseases in the Reserve, e.g., TB, STDs,\ldots 
  \item
    The data is not necessarily representative and we had to use educated guesses
    for mixing parameters.
  \item
    Have not yet considered more complex epidemiological and behavioral
    dynamics. For example, group structure can affect the coupled contagion of
    disease and behaviors that could encourage or impede the spread of
    disease~\autocite{SmaldinoJones2021}. Our compartmental model
    incorporated group structure and contact rate heterogeneity, but finer
    details of the social network structure can induce complex
    epidemiological dynamics~\autocite{Arthur2017}. In the case of Dourados
    and possibly other Indenous communities is the tendency for several
    generations of families to live in close proximinty to one another,
    often communally with other families as well, making social isolation
    culturally unpalatable and, consequentially, unfeasible. Adaptation to
    evolving, uncertain epidemiological social and environmental contexts is
    one further factor that can complicate epidemiological
    dynamics~\autocite{Arthur2021} and willingness to learn beneficial public
    health behaviors from others~\autocite{Turner2023a}.
  \item
    Currently we do not connect how much it costs to implement these programs,
    and there may be difficult questions over what equality means: is it
    spending the same amount of money per group or preventing the same number of
    hypothetical infections for each group? Of course a scientific article 
    cannot provide the answer, but here we do provide a program for evaluating
    different choices to prototype planned responses to pandemic outbreaks and
    reason quantitatively about alternative courses of action.
\end{itemize}

>>>>

To quantitatively compare hypothetical intervention strategies we will analyze 
model outcomes in different contexts, modeled as different parameter settings.
Since this document is for planning the study and analysis, here we present a sketch
of results we plan to obtain and subsequent analysis of these results. In the first
analysis, we examine how intervention efficacy varies with the possibility of
re-infection and mobility reduction strategy
(Figure~\ref{fig:ByReinfectionByStrategy}). In the second, we examine how
intervention efficacy varies for different groups for a given mobility reduction
strategy (Figure~\ref{fig:ByGroupReduction_GlobalStrategy}). These analyses 
promise to account for some important sources of complexity in 
the Dourados Reserve-city system that could affect intervention outcomes.

\subsection{Aid strategy and epidemiological details affect incidence reduction}

We hypothesize that reducing incidence of diseases or variants with 
re-infection (e.g., Zeta COVID variant) may require greater investment 
for the same reduction in incidence
compared to diseases or variants without re-infection (e.g., B.1.1 variant).
This would result in a sharper s-curve from zero reduction 
to maximum reduction, with 50\% of maximum reduction occurring 
at a greater value of $\Delta \gamma$, across aid strategies (illustrated in
difference between orange, re-infection, and blue lines, no re-infection, in 
Figure~\ref{fig:ByReinfectionByStrategy}).

We also hypothesize that it will be more effective to target individuals based on
geographic location compared to ethnic identity, assuming there are no significant
differences in transmissibility, which may exist due to physical, medical or
behavioral differences between ethnic groups \mt{we plan to return to this point in
the Discussion}. This will result in efficacy curves that reach maximal
effectiveness sooner (compare \emph{80-20} strategy with 
80\% to Jaguapiru–dashed lines–to \emph{80-20}, with 80\% to Guarani (all)–dotted
lines–and the \emph{Equal} strategy–solid line;
Figure~\ref{fig:ByReinfectionByStrategy}).

@@

\subsection{Intervention efficacy varies by group}

@@

\begin{itemize}
  \item 
    Application to other problematic diseases in the Reserve, e.g., TB, STDs,\ldots 
  \item
    The data is not necessarily representative and we had to use educated guesses
    for mixing parameters.
  \item
    Have not yet considered more complex epidemiological and behavioral
    dynamics. For example, group structure can affect the coupled contagion of
    disease and behaviors that could encourage or impede the spread of
    disease~\autocite{SmaldinoJones2021}. Our compartmental model
    incorporated group structure and contact rate heterogeneity, but finer
    details of the social network structure can induce complex
    epidemiological dynamics~\autocite{Arthur2017}. In the case of Dourados
    and possibly other Indenous communities is the tendency for several
    generations of families to live in close proximinty to one another,
    often communally with other families as well, making social isolation
    culturally unpalatable and, consequentially, unfeasible. Adaptation to
    evolving, uncertain epidemiological social and environmental contexts is
    one further factor that can complicate epidemiological
    dynamics~\autocite{Arthur2021} and willingness to learn beneficial public
    health behaviors from others~\autocite{Turner2023a}.
  \item
    Currently we do not connect how much it costs to implement these programs,
    and there may be difficult questions over what equality means: is it
    spending the same amount of money per group or preventing the same number of
    hypothetical infections for each group? Should past discrimination and
    government-imposed hardship be included in the calculus of developing
    mobility-reduction strategies? Of course a scientific article 
    cannot provide the answer, but here we do provide a program for evaluating
    different choices to prototype planned responses to pandemic outbreaks and
    reason quantitatively about alternative courses of action.
\end{itemize}
